Preparing your manuscript

### 2025-02-21 article/article.qmd
notes about grad temp

data <- purrr::map(Sys.glob(here("data/eke_mean/*hist*mean.nc")), function(f) {

@@

  if (meta[[1]][["var"]] == "eke") {
    var <- "ua"
  } else {
    var <- "tasmax"
  }
  file_scenario <- here(paste0("data/eke_mean/", meta[[1]][["var"]], "_", meta[[1]][["model"]], "_ssp585_mean.nc"))

@@

In this paper we build on the drops in temperature methodology to evaluate the change in frequency of cold fronts in a warmer future compared with the recent past using a set of models from CMIP6. In an effort to understand possible causes to the potential changes, we explored the relationship between changes in front frequency and transient eddy kinetic energy, that is directly related to storm tracks. Several studies have shown a poleward shift on the storm tracks in future climate change scenarios, particularly in the Southern Hemisphere [@ogorman2010; @tamarin2017; @chemke2022], and these changes could potentially influence the location and intensity of future cold fronts.

@@

@@

In the Southern Hemisphere over land, the change is generally positive with 5 events per year in some areas. Other authors have also found an increase in the frequency of fronts in the south and east regions of Africa and Australia [@catto2014; @konstali2024]. @fig-trend a and b) show over the southern oceans a decrease in the frequency of $\Delta T$ between 30$^{\circ}$ and 45$^{\circ}$ S (depending on the longitude) and an increase to the south. This pattern indicates that the areas of higher front frequency will shift poleward in the SSP5-8.5.

**Increase over land vs decrease over sea could be related also to the thermal inertia / land get hotter?**

@@

**why? **

@@

@fig-eke explores the relationship between the change in EKE and frequency of $\Delta T$ comparing the scenario SSP5-8.5 (2080-2100) and the recent past (1980-2000). In the Northern Hemisphere, both EKE and the frequency of $\Delta T$ show mainly negative changes (@fig-eke a). The zonal mean shows that this positive relationship holds well over oceans with negative changes of $\Delta T$ frequency and EKE around 40$^{\circ}$ N (@fig-eke b). In the Southern Hemisphere, the overall relationship appears negative, primarily due to increasing $\Delta T$ frequency over land while EKE maxima shifts poleward. **why? are we detecting changes in temp not related to cyclones and frontal things?** Over the sea, the changes are consistent with a poleward shift of the storm tracks (already reported by @ogorman2010, @tamarin2017, @chemke2022) and the maximum frequency of $\Delta T$.

>>>>

data <- purrr::map(c(Sys.glob(here("data/eke_mean/*hist*mean.nc")), Sys.glob(here("data/deltat/*hist*mean.nc"))), function(f) {

@@

  if (meta[[1]][["var"]] == "eke") {
    var <- "ua"
  } else {
    var <- "tasmax"
  }
  file_scenario <- Sys.glob(here(paste0("data/", c("eke_mean", "deltat"), "/", meta[[1]][["var"]], "_", meta[[1]][["model"]], "_ssp585_mean.nc")))

@@

In this paper we build on the drops in temperature methodology to evaluate the change in frequency of cold fronts in a warmer future compared with the recent past using a set of models from CMIP6. In an effort to understand possible causes to the potential changes, we first studied the response of the meridional temperature gradient and it relationship with the change in front frequency. We explored the relationship between changes in front frequency and transient eddy kinetic energy, that is directly related to storm tracks. Several studies have shown a poleward shift on the storm tracks in future climate change scenarios, particularly in the Southern Hemisphere [@ogorman2010; @tamarin2017; @chemke2022], and these changes could also influence the location and intensity of future cold fronts.

@@

### Meridional temperature gradient

We calculated the meridional temperature gradient ($\partial T/ \partial y$) at each grid point using central finite differences for the temperature at 850 hPa. We then area-averaged $\partial T/ \partial y$ over different regions, i.e.  between 25$^{\circ}$ and 65$^{\circ}$ S. This process produces the same results as if we calculate the temperature difference between2 latitudes but give us the advantage of calculating the mean $\partial T/ \partial y$ for regions over land and sea independently.

@@

In the Southern Hemisphere over land, the change is generally positive with 5 events per year in some areas. Other authors have also found an increase in the frequency of fronts in the south and east regions of Africa and Australia [@catto2014; @konstali2024]. @fig-trend a and b) show over the southern oceans a decrease in the frequency of $\Delta T$ between 30$^{\circ}$ and 45$^{\circ}$ S (depending on the longitude) and an increase to the south. This pattern indicates that the areas of higher front frequency will shift poleward in the SSP5-8.5. This region is also associated with an increase in $\partial T/ \partial y$ in storm track activity (not show **or supplemental material**).

@@

@@

@fig-eke explores the relationship between the change in the frequency of $\Delta T$ and the response of $\partial T / \partial y $ and EKE by comparing the scenario SSP5-8.5 (2080-2100) and the recent past (1980-2000). There is a positive relationship between the changes in $\Delta T$ and $\partial T / \partial y$ in both hemispheres. The reduction in the frequency of fronts in the Northern Hemisphere could be associated to the decrease of $\partial T / \partial y$ in low levels already shown by other authors [e.g. @catto2014] while the opposite is true for the Southern Hemisphere (@fig-eke a). 
We see a similar relationship between changes in $\Delta T$ and EKE except for the Southern Hemisphere over land. The reason for this it is difference could be associated to many factors. 
The change of the stormtracks in the SH is mainly associated to a poleward shift and may not have an effect on frequency of fronts over land the same way the $\partial T / \partial y$ does. @harvey2013 found that the stormtrack response in the SH during summer is associated to the change in the low level $\partial T / \partial y$ mainly bellow 65$^{\\circ}$ S, i.e. over the southern oceans.
